<!--
id: "doe-meet-bedroom"
created: "Sat Jan 31 11:49:48 2026"
attribution: DTK
use: Chap 8 drill
status: "ready-to-use"
chapter: 13
mode: "outline"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-doe-meet-bedroom}
This set of problems involves a linear combination of two core modeling functions, but there is an added twist: One or both of the functions has been input scaled to move it to the right or the left.

When you have successfully reconstructed the function displayed, cut-and-paste the line `newf <- ...` 

```{webr}
#| echo: false
#| autorun: true



```


```{webr}
#| caption: Plot a linear combination
# Change the values of the coefficients and the function names as appropriate
newf <- function(x) osc(x) + 2 * hillside(x - 2)

# You can use this line as is, no changes needed
slice_plot(newf(x) ~ x, domain(x = -5:5))
```

-----

:::: {.callout-note}
## Functions to reconstruct

::: {.panel-tabset}

## A.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-auzgx
newf <- function(x) hillside(x + 2) + 0.1 * same(x)
slice_plot(newf(x) ~ x, domain(x = -5:5), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-A-3ks")`

::: {.callout-tip collapse=true}
## Hints:

One of the functions is `hillside()`. 
:::

::: {.callout-tip collapse=true}
## Answer
[`hillside(x + 2) + 0.1 * same(x)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: TYf-yds"}
:::


## B.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-lkhze
newf <- function(x) hillside(x + 3) + 2 * hillside(x - 3)
slice_plot(newf(x) ~ x, domain(x = -5:5), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-B-8473")`

::: {.callout-tip collapse=true}
## Hint:
Both of the functions are `hillside()`. 

:::

::: {.callout-tip collapse=true}
## Answer
[`hillside(x + 3) + 2 * hillside(x - 3)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: GOf-2wS"}
:::

## C. 

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-yuceq
newf <- function(x) -hill(x + 2) + 2 * hill(x - 1)
slice_plot(newf(x) ~ x, domain(x = -5:5), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-C-d92")`

::: {.callout-tip collapse=true}
## Hint:
One goes down, one goes up.

:::

::: {.callout-tip collapse=true}
## Answer
[`hill(x + 2) - 2 * hill(x - 1)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: KID-2wS"}
:::

:::

::::

:::
 <!-- end of exr-doe-meet-bedroom -->
