<!--
id: "doe-meet-kitchen"
created: "Sat Jan 31 11:49:48 2026"
attribution: TBA
status: "ready-to-use"
chapter: 13
mode: "outline"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-doe-meet-kitchen}
The R chunk contains a computer expression to create and plot a linear combination of two functions `osc(x)` and `same(x)`, each multiplied by a numerical coefficient. Use this as a template to re-create the requested plots. Keep in mind that you need to replace `osc()` and `same()` with core modeling functions appropriate for the task.

When you have successfully reconstructed the function displayed, cut-and-paste the line `newf <- ...` 

```{webr}
#| echo: false
#| autorun: true



```


```{webr}
#| caption: Plot a linear combination
# Change the values of the coefficients and the function names as appropriate
newf <- function(x) 2 * osc(x) - 3 * same(x)

# You can use this line as is, no changes needed
slice_plot(newf(x) ~ x, domain(x = -4:4))
```

-----

:::: {.callout-note}
## Functions to reconstruct

::: {.panel-tabset}

## A.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-denfb
newf <- function(x) osc(x) - 2 * hillside(x)
slice_plot(newf(x) ~ x, domain(x = -4:4), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-A-d7w")`

::: {.callout-tip collapse=true}
## Hints:

1. One of the functions is `hillside()`. 
2. The half-way up level for the other function is at 0 to the left and -2 to the right.
:::

::: {.callout-tip collapse=true}
## Answer
[`1 * osc - 2 * hillside(x)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: Dcf-2wS"}
:::


## B.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-ukzop
newf <- function(x) 0.5 * hillside(x) + 3 * hill(x)
slice_plot(newf(x) ~ x, domain(x = -4:4), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-B-x7d")`

::: {.callout-tip collapse=true}
## Hint:
One of the functions is `hillside()`. 

:::

::: {.callout-tip collapse=true}
## Answer
[`0.5 * hillside(x) + 3 * hill(x)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: Gsf-2wS"}
:::

## C.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-riysx
newf <- function(x) 4 * flat(x) + 3 * osc(x)
slice_plot(newf(x) ~ x, domain(x = -4:4), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-C-klwd")`

::: {.callout-tip collapse=true}
## Hint:
Obviously, one of the function is `osc()`. But on its own, the output of `osc()` is bounded by -1 and 1.

:::

::: {.callout-tip collapse=true}
## Answer
[`4 * flat(x) + 3 * osc(x)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: JEf-2wS"}
:::

## D.

Reconstruct the function graphed below as a linear combination of two core modeling functions.
```{r}
#| echo: false
#| label: C13-uateu
newf <- function(x) 1*exp2(x) - 3*same(x)
slice_plot(newf(x)~ x, domain(x = -4:4), npts = 500) |>
  gf_theme(theme_minimal(base_size = 14))
```

`r devoirs::devoirs_text("dmk-D-sws")`

::: {.callout-tip collapse=true}
## Hint:
On the left side, the output is more or less a straight line, but sloping downward. `same()` is our straight-line core modeling function, but ordinarily it slopes upward.

:::

::: {.callout-tip collapse=true}
## Answer
[`1 * exp2(x) - 3 * same(x)`]{.shortanswer `r devoirs:::answer_style()` data-tooltip="Ans id: JEf-2wS"}
:::




:::

::::

:::
 <!-- end of exr-doe-meet-kitchen -->
