<!--
id: "fox-tell-dress"
created: "Fri Feb 20 09:33:20 2026"
attribution: TBA
status: "ready-to-use"
chapter: 13
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-fox-tell-dress}

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: C13-hzqdf
dyn1 <- makeODE(dy ~ 2, dx ~ 2-y)
sol1a <- integrateODE(dyn1, domain(t=0:2), x = 0, y =0)
sol1b <- integrateODE(dyn1, domain(t=0:2), x = 2, y =1)

dyn2 <- makeODE(dy ~ x, dx ~ x + 2-y)
sol2a <- integrateODE(dyn2, domain(t=0:1.6), x = 1, y =1)
sol2b <- integrateODE(dyn2, domain(t=0:.6), x = 2, y =0)
```

```{r}
#| echo: false
#| label: C13-cqrtk
flow_field(dyn1, domain(x = 0:5, y = 0:5)) |>
  traj_plot(x(t) ~ y(t), sol1a, color = "blue", nt=0, inherit=FALSE, linewidth = 2, alpha = 0.5) |>
  traj_plot(x(t) ~ y(t), sol1b, color = "green", nt=0, inherit=FALSE, linewidth = 2, alpha = 0.5) |>
  traj_plot(x(t) ~ y(t), sol2a, color = "tomato", nt=0, inherit=FALSE, linewidth = 2, alpha = 0.5) |>
  traj_plot(x(t) ~ y(t), sol2b, color = "gray", nt=0, inherit=FALSE, linewidth = 2, alpha = 0.75) |>
  gf_refine(coord_fixed()) |>
  gf_theme(theme_minimal(base_size = 16))
```

1. `r devoirs::devoirsTF(FALSE, id = "traj-ftd-1-lx")` For the flow shown above, [red]{style="color: red;"} is a trajectory.

2. `r devoirs::devoirsTF(TRUE, id = "traj-ftd-2-l7y")` For the flow shown above, [green]{style="color: green;"} is a trajectory.

3. `r devoirs::devoirsTF(FALSE, id = "traj-ftd-3-kw")` For the flow shown above, [red]{style="color: red;"} is a trajectory.

4. `r devoirs::devoirsTF(FALSE, id = "traj-ftd-3-luy")` For the flow shown above, [gray]{style="color: darkgray;"} is a trajectory.

```{r}
#| echo: false
#| label: C13-pypey
traj_plot(x(t) ~ t, sol1b) |> 
  gf_labs(x = "t", y = "??", subtitle="Time series 1")
```

5. Time series 1 matches which component of which trajectory?

    a. [gray-x]{style="color: darkgray;"}
    b. [gray-y]{style="color: darkgray;"}
    c. [red-x]{style="color: red;"}
    d. [red-y]{style="color: red;"}
    e. [green-x]{style="color: green;"}
    f. [green-y]{style="color: green;"}
    g. [blue-x]{style="color: blue;"}
    h. [blue-y]{style="color: blue;"}

```{mcq}
#| label: ts-ftd-5-wkd
#| inline: true
1. a
2. b
3. c
4. d
5. e  [correct]
6. f
7. g
8. h
9. None of the others.
```


```{r}
#| echo: false
#| label: C13-khody
traj_plot(x(t) ~ t, sol2a) |> 
  gf_labs(x = "t", y = "??", subtitle="Time series 2")
```

6. Time series 2 matches which component of which trajectory?

    a. [gray-x]{style="color: darkgray;"}
    b. [gray-y]{style="color: darkgray;"}
    c. [red-x]{style="color: red;"}
    d. [red-y]{style="color: red;"}
    e. [green-x]{style="color: green;"}
    f. [green-y]{style="color: green;"}
    g. [blue-x]{style="color: blue;"}
    h. [blue-y]{style="color: blue;"}

```{mcq}
#| label: ts-ftd-6-wkd
#| inline: true
1. a
2. b
3. c [correct]
4. d
5. e 
6. f
7. g
8. h
9. None of the others.
```

:::
 <!-- end of exr-fox-tell-dress -->
