<!--
id: "lion-let-stair"
created: "Wed May 13 08:00:41 2026"
attribution: DTK
chapter: 10
status: sketch
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-lion-let-stair}
TURN THIS INTO EXERCISES, giving both a function graph and a graph of the gradient field.

[Whenever you can plot a function]{style="font-variant:small-caps;"} finding a maximum is easy.A glance at the graph will tell you which input has the highest output.

As an example, here is a randomly selected function with one input.

::: {.content-hidden unless-format="html"}
```{webr}
#| caption: Optimize this!
h <- doodle_fun( ~ x, seed = 832)
slice_plot(h(x) ~ x, 
           domain(x = -5:5)) 
```
::::

::: {.content-hidden unless-format="pdf"}
```{r}
#| echo: false
#| label: C10-lgobq
h <- doodle_fun( ~ x, seed = 832)
slice_plot(h(x) ~ x, 
           domain(x = -5:5)) 
```
::::

You can locate the maximum (or the minimum) almost instantly.

Another example of finding minima, now for a function of two inputs. Try this randomly generated function:

:::: {.content-hidden unless-format="html"}
```{webr}
#| caption: Find the minimum
g <- doodle_fun(~ x + y, seed = 101)
contour_plot(g(x, y) ~ x & y, 
             domain(x = -4:4, y = -4:4))
```

Can you see where the minimum is? Basic contour-graph reading skills will lead you there; look for the small circularly-shaped contour that has a lower output level than the surrounding circularly-shaped contours. By changing the number after seed =  to another integer, you can generate another random function. You’ll find it easy to locate the minimum (which might be on the border).
::::

:::: {.content-hidden unless-format="html"}
```{r}
#| echo: false
#| label: C10-qnjmw
g <- doodle_fun(~ x + y, seed = 101)
contour_plot(g(x, y) ~ x & y, 
             domain(x = -4:4, y = -4:4))
```

Can you see where the minimum is? Basic contour-graph reading skills will lead you there; look for the small circularly-shaped contour that has a lower output level than the surrounding circularly-shaped contours.
::::
:::


:::
 <!-- end of exr-lion-let-stair -->
