<!--
id: "calf-catch-spoon"
created: "Mon Aug 24 13:23:18 2026"
attribution: DTK
uses: Activity about functions
status: "sketch"
chapter: 6
mode: "outline"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-calf-catch-spoon}
::: {.callout-note}
## Using functions "inside-out"

Someone shopping for a house needs to know what price is affordable to them. Typically, people have a good idea of how much they can pay monthly, but need to translate this into a new form: the *principal* of the loan what will generate such a payment.

For this purpose, the `monthly_payment()` function is set up backwards. To use `monthly_payment()`, you need to know the principal---it will calculate the corresponding monthly payment. 

Even so, `monthly_payment()` can lead you to the answer. You just have to turn it inside out. This involves a simple procedure. Say you can afford a $2500 monthly payment. What is the corresponding principal?

Start by guessing. Could you afford a million-dollar loan? Let's see:

```{r echo=FALSE}
library(QRA)
```

```{r}
monthly_payment(1e6, 6, 30)
```

No good. The million-dollar loan would cost about $6000 per month. If you completely understood the mechanics of the monthly-payment calculation, you could use algebra to find what principal you can afford.

But you don't need to understand the function, you just need to be able to use it. Since one million dollars is too rich for your situation, try half-a-million dollars.

```{r}
monthly_payment(0.5e6, 6, 30)
```

Still too much! Keep on trying until you hit the spot.



:::

:::
 <!-- end of exr-calf-catch-spoon -->
