<!--
id: "oak-form-futon"
created: "Tue Sep 30 12:54:39 2025"
attribution: TBA
status: "draft"
chapter: 8
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-oak-form-futon}
## Example: Predicting the US population


Every 10 years, starting in 1790, the US Census Bureau carries out a constitutionally mandated census: a count of the current population. The overall count as a function of year is shown in `r xf("fig-pop-graph")`. [[Source](https://en.wikipedia.org/wiki/Demographic_history_of_the_United_States)
]

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: C08-etbpq
USPop <- readr::read_csv("www/census-totals.csv")
Tmp <- USPop |> 
  mutate(population=prettyNum(population, big.mark=",")) %>%
  select(-growth_rate)
Table <- if (knitr::is_html_output()) {
  DT::datatable(Tmp) 
} else {
  knitr::kable(Tmp)
}
USPop <- USPop |> 
  mutate(population = population/1e6)
```

In the 230 years spanned by the census data, the US population has grown 100-fold, from about 4 million in 1790 to about 330,000,000 in 2020.

::: {#fig-pop-graph2}
```{r echo=FALSE, warning=FALSE} 
#| layout-ncol: 2
#| fig-subcap: 
#| - Linear scale
#| - Semi-log scale
#| label: C08-teesb
popfun <- spliner(population ~ year, data = USPop)
USPop <- USPop %>%
  mutate(yearly_growth_rate = 100*((1+growth_rate/100)^(1/10) - 1))
mod <- lm(population ~ poly(year, 2), data = USPop)
modf <- makeFun(mod)
P1 <-gf_point(population ~ year, data = USPop) %>%
  gf_line(population ~ year) %>%
  gf_labs(y="Population (millions)") %>%
  gf_lims(x = c(1790, 2100), y=c(0,500))
P2 <- gf_point(population ~ year, data = USPop) %>%
  gf_point(500 ~ 2100, alpha=0) %>%
  gf_line(population ~ year) %>%
  gf_labs(y="Population (millions)") %>%
  gf_refine(scale_y_log10(breaks=100*(0:5))) %>%
  gf_lims(x = c(1790, 2100))
P1
P2
```

US population since 1790
:::



It is tempting to look for simple patterns in such data. Perhaps the US population has been growing exponentially. A semi-log plot of the same data suggests that the growth is only very roughly exponential. A truly exponential process would present as a curve with a constant derivative, but the derivative of the function in the graph is decreasing over the centuries. 

Insofar as the slope over the semi-log graph is informative, it amounts to this quantity:
$$\partial_t \ln(\text{pop}(t)) = \frac{\partial_t\, \text{pop}(t)}{\text{pop}}$$
This is the *per-capita* rate of growth, that is, the rate of change in the population divided by the population. Conventionally, this fraction is presented as a percentage: percentage growth in the population per year, as in @fig-pop-growth-2.

::: {#fig-pop-growth-2}
```{r echo=FALSE, warning=FALSE}
#| label: C08-hicfd
growth_mod <- lm(yearly_growth_rate ~ year, data = USPop)
growth_fun <- makeFun(growth_mod)
gf_point(yearly_growth_rate ~ year, data = USPop) %>%
  gf_lims(x = c(1790, 2100), y=c(-.5,4)) %>%
  gf_lm() %>%
  gf_labs(y = "% growth per year") 
```

Annual per capita growth rate of the US population (percent)
:::

The dots in the graph are a direct calculation from the census data.  There is a lot of fluctuation, but an overall trend stands out: the population growth rate has been declining since the mid-to late 1800s. The deviations from the trend are telling and correspond to historical events. There is a relatively low growth rate seen from 1860 to 1870: that is  the effect of the US Civil War. The Great depression is seen in the very low growth from 1930 to 1940. Baby Boom: look at the growth from 1950-1960. The bump from 1990 to 2000? Not coincidentally, the 1990 Immigration Act substantially increased the yearly rate of immigration. 

If the trend in the growth rate continues, the US will reach zero net growth about 2070, then continue with negative growth. Of course, negative growth is just decline. A simple prediction from @fig-pop-growth is that the argmax of the US population---that is, the year that the growth rate reaches zero---will occur around 2070.

How large will the population be when it reaches its maximum? 

In Block 2, we dealt with situations where we know the function $f(t)$ and want to find the rate of change $\partial_t f(t)$. Here, we know the rate of change of the population and we need to figure out the population itself, in other words to figure out from a known $\partial_t f(t)$ what is the unknown function $f(t)$.  

The process of figuring out $f(t) \longrightarrow \partial_t f(t)$ is, of course, called ***differentiation***. The opposite process, $\partial_t f(t) \longrightarrow f(t)$ is called ***anti-differentiation***. 

In this block we will explore the methods for calculating anti-derivatives and some of the settings in which anti-derivative problems arrive. 

::: {#thm-US-pop-growth style="display: none;"}
---Growth at a continuously decreasing growth rate.
:::
::: {.callout-note icon=false data-latex=""} 
## @thm-US-pop-growth Population growth

The predictions from the accumulate-population-growth model are shown as a [magenta]{style="color: magenta;"} line in @fig-pop-prediction-bad.

::: {#fig-pop-prediction-bad}
```{r echo=FALSE, warning=FALSE}
#| label: C08-oofwc
growth_mod <- lm(yearly_growth_rate ~ year, data = USPop)
growth_fun <- makeFun(growth_mod)
accum <- tibble(year = 2021:2130,
                pop = 331 * cumprod(1 + growth_fun(2021:2130)/100))

gf_point(population ~ year, data = USPop) %>%
  gf_labs(y="Population (millions)") %>%
  gf_lims(x = c(1790, 2130), y=c(0,450)) %>%
  gf_line(pop ~ year, data = accum, color="magenta") %>%
  gf_point(404.4 ~ 2060, color="dodgerblue", size=1)
```

Predicted US population based on the historical linear decline in per-capita growth.
:::

According to the accumulation model, the population peaks in 2075 at 390 million. We will be back down to the present population level in about 100 years. 

Professional demographers make much more sophisticated models using detailed data from many sources. The demographers at the US Census Bureau predict that the population will reach a maximum of 404 million in 2060, shown by the little blue dot in @fig-pop-prediction-bad.  that is not too different from what we got by analyzing just the raw census numbers. 
:::

:::
 <!-- end of exr-oak-form-futon -->
