<!--
id: "kitten-close-coat"
created: "Tue Jan 20 14:02:46 2026"
attribution: TBA
status: "ready-to-use"
chapter: 5
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-kitten-close-coat}
Home chefs are advised that there is a safe temperature for cooking poultry: a thermometer inserted into the deepest part of the flesh should reach 165 deg F. This often leads to dry meat. In contrast, "rotisserie chicken," the sort of fresh, pre-cooked chicken sold by grocery stores, is almost always moist and tasty. How does the rotisserie method achieve this? 

There is actually a wide range of temperatures at which chicken can be cooked safely. The key is to hold the meat at that temperature for long enough. At 165 deg F, safety is achieved *instantly*.  
@fig-safe-poultry-temp shows the required holding time at a range of temperatures:


```{r}
#| echo: false
#| label: C05-wdzbe
Poultry <- tibble::tribble(
~temp, ~c_time, ~t_time,
136, 81.4, 70.8,
137, 65.5, 58.5,
138, 52.9, 48.5,
139, 43, 40.4,
140, 35, 33.7,
141, 28.7, 28.2,
142, 23.5, 23.7,
143, 19.3, 19.8,
144, 15.9, 16.6,
145, 13, 13.8,
146, 10.6, 11.5,
147, 8.6, 9.4,
148, 6.8, 7.7,
149, 5.4, 6.2,
150, 4.2, 4.9,
151, 3.1, 3.8,
152, 2.3, 2.8,
153, 1.6, 2.1,
154, 1.1, 1.6,
155, 0.9, 1.3,
156, 0.72, 1,
157, 0.57, 0.84,
158, 0.45, 0.68,
159, 0.36, 0.55,
160, 0.28, 0.45,
161, 0.22, 0.36,
162, 0.18, 0.30,
163, 0.14, 0.24,
164, 0.10, 0.19,
165, 0.05, 0.09,
)
```

```{r}
#| label: fig-safe-poultry-temp
Poultry |>
  gf_line(c_time ~ temp) |>
  gf_labs(x = "Safe temperature (deg F)", y = "Holding time") |>
  gf_theme(theme_minimal())
```
:::

1. Suppose you determine to cook your chicken at 140 deg F. Assuming it takes two hours of cooking to ensure that the entire chicken reaches that temperature, how much longer do you need to cook the chicken in order to make it safe?

```{mcq}
#| label: kcc-1-kes
#| inline: true
1. 25 minutes
2. 30 minutes
3. 35 minutes [correct]
4. 40 minutes
5. No such time.
```

Naturally, you want to avoid any risk of food poisoning. This could happen, for instance, if your oven runs cool. For instance, an oven set to 140 deg F might actually have cold spots at, say, 135 deg F.

There are two ways to guard against this. One is simply to hold the chicken at the cooking temperature for longer than the answer in Question (1). 

2. If you want to guard against the possibility of an oven temperature of 135 deg F, how long do you need to hold the chicken at cooking temperature for safety? (Remember, this is the time *beyond* the original two hours needed to get the bird up to the cooking temperature.) 

```{mcq}
#| label: kcc-2-kes
#| inline: true
1. 40 minutes
2. 60 minutes
3. 80 minutes 
4. No such time shown in graph. [correct]
```

Home chefs are used to a cooking method where the temperature is set very high (e.g. 350 deg F) and the time carefully regulated to avoid overcooking. Better results can be had at much lower temperatures so long as you take care to hold the bird at the temperature for long enough and check and calibrate your oven to ensure that it is at the right temperature. Commercial rotisseries have ovens set up for this purpose.



 <!-- end of exr-kitten-close-coat -->
