<!--
id: "kid-ring-screen"
created: "Mon Nov  3 11:10:56 2025"
attribution: TBA
status: "ready-to-use"
chapter: 11
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}



::: {#exr-kid-ring-screen}
The figure shows six different instantaneous configurations for a pendulum as it swings. The curved arrow shows the velocity of the bob. The length of the curved arrow is the velocity times one second.

```{r}
#| echo: false
#| warning: false
#| label: fig-pendulums-6
#| layout-ncol: 3
#| fig-cap: " "
#| fig-subcap:
#| - " "
#| - " "
source("_software/pendulum.R")
draw_pendulum1(pi/8, -0.2)
draw_pendulum1(-pi/4, 0.05)
draw_pendulum1(pi/7, -0.3)
draw_pendulum1(-pi/6, 0.5)
draw_pendulum1(pi/10, -0.1)
draw_pendulum1(-pi/10, +1)
```

1. Which frame shows the pendulum where it is near to the topmost point on its swing.
```{mcq}
#| label: krs-topmost
#| inline: true
1. a
2. b [correct]
3. c
4. d
5. e
6. f
```

2. For drawing purposes, @fig-pendulums6-state-space gives you an empty image of the state space for the pendulum.

```{r}
#| label: fig-pendulums6-state-space
gf_blank(0 ~ 0) |>
  gf_lims(y = c(-1,1)) |>
  gf_labs(y = "velocity", x = (expression(paste(theta, " (deg)")))) |>
  gf_refine(
    scale_x_continuous(
      limits = c(-100,100), 
      breaks = c(-90,-60, -45, -30, -15, 0, 15, 30, 60, 45, 60, 90))) |> 
  gf_theme(theme_minimal())
```

Your task: For each of the six pendulum pictures in `r xf("fig-pendulums-6")`, write the corresponding letter a-f in the right position in the state space. Imagine that the fastest velocity in `r xf("fig-pendulums-6")` is 1. Your answers can be approximate, but they should be correct order relative to one another.

3. Return to `r xf("fig-pendulums6-state-space")`, adding in a vector showing the rate of change of state for each picture with respect to time. Hint: You can tell the rate of change of $\theta$ by referring to the arrows in `r xf("fig-pendulums-6")`. But for the rate of change with respect to $\theta$, you'll have to draw on your intuition about how a pendulum works. Your answers will necessarily be approximate, but make sure to get the *sign* right for each quantity.

:::
 <!-- end of exr-kid-ring-screen -->
