<!--
id: "snail-sit-saucer"
created: "Tue Dec  9 16:41:39 2025"
attribution: TBA
status: "ready-to-use"
chapter: 11
mode: "discussion"
-->

{{< include ../exercise-header.qmd >}}



::: {#prp-snail-sit-saucer}
In-class activity.

Give each student a piece of paper with several number lines drawn on it. The number lines should have ticks, but no numerical labels.

1. For an in-class discussion, use a simulation to display values one at a time of a random variable. Showing 20 of these to the students, and having them mark each value on a number line, ask the students to sketch a relative probability distribution for the random quantity.

2. For a discussion about parameters for the distributions, ask them to estimate the parameters of the uniform distribution that capture numbers secretly generated, one at a time, from a uniform distribution. They should mark the successive numbers on one of their number lines. Have them remark each time a new random number causes them to revise their estimate of the max and min parameters.

```{r}
#| label: C11-chwlt
# don't let them see this.
process1 <- function() {runif(1, min = 2, max = 10)}
process1() # repeat many times 
```

3. Very much like (2), put use `process2()` to generate the numbers.

```{r}
#| label: C11-dboaw
# don't let them see this.
process2 <- function() 4 + rexp(1, rate=1/3)
process2()
```

Since the numbers are not actually from a uniform distribution (but don't tell them this!) the estimate of the max parameter will not be stable. Generate a score of values, so that at least one is pretty much an outlier compared to the others. 

Ask whether they think we should reconsider the uniform distribution as the model for `process2()`. What makes them think that a uniform distribution in unlikely to be a good description of these data. (We are leading up to the concept of "likelihood." The problem with the uniform distribution for the `process2()` data is that it leaves a lot of the probability mass over parts of the domain that are empty.)

:::
 <!-- end of prp-snail-sit-saucer -->
