<!--
id: "fly-hide-mattress"
created: "Sat Dec 20 08:53:02 2025"
attribution: TBA
status: "draft"
chapter: 7
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}


::: {#exr-fly-hide-mattress}
AI generated question: When navigating a river, your speed is affected by the current, wind, and your own physical stamina. To model this, we can use a relative probability function to visualize the range of possible outcomes.

Scenario: You are planning a canoe trip 8 miles upstream to a campsite. Under perfect conditions, you paddle at a constant rate of 4 miles per hour (mph). However, the river has a variable current that flows against you.

@fig-fhm-speed shows your professor's sense of the relative probability for your  "Net Upstream Speed" (your paddling speed minus the river's current). 

```{r}
#| label: fig-fhm-speed
#| echo: false
#| fig-cap: "Relative probability for net speed."

# --- 1. Define Parameters ---
distance <- 8 # miles

# --- 2. Create Data for Speed Uncertainty ---
# We define a triangular distribution where speed is between 1 and 3 mph, 
# with the peak (mode) at 2 mph.
speed_df <- data.frame(v = seq(1, 3, length.out = 500))
speed_df$likelihood <- ifelse(speed_df$v <= 2, 
                               speed_df$v - 1,   # Upslope
                               3 - speed_df$v)   # Downslope

# --- 3. Create Data for Time Uncertainty ---
# Transformation: T = distance / V
# Math: g(t) = f_v(distance/t) * | -distance / t^2 |
time_df <- data.frame(t = seq(distance/3, distance/1, length.out = 500))

# Calculate the corresponding speed for each time point
time_df$v_equiv <- distance / time_df$t

# Calculate the likelihood of that speed
time_df$f_v <- ifelse(time_df$v_equiv <= 2, 
                      time_df$v_equiv - 1, 
                      3 - time_df$v_equiv)

triangle <- function(t) {
  ifelse(t < 2.67, 0,
         ifelse(t < 5, (t - 2.67),
                ifelse(t < 8, (8 - t)*2.33/3 , 0)
                )
  )/5.33
}

# Apply the Jacobian adjustment (distance / t^2) to keep the area under the curve = 1
time_df$likelihood <- time_df$f_v * (distance / (time_df$t^2))
time_df$triangle <- triangle(time_df$t)

# --- 4. Plotting ---

# Speed Plot
p1 <- ggplot(speed_df, aes(x = v, y = likelihood)) +
  geom_area(fill = "skyblue", alpha = 0.5, color = "blue") +
  labs(title = "Uncertainty in Speed (V)",
       subtitle = "Symmetrical Triangular Distribution",
       x = "Net Speed (mph)",
       y = "Probability") +
  theme_minimal()

# Time Plot
p2 <- ggplot(time_df, aes(x = t, y = likelihood)) +
  geom_area(fill = "salmon", alpha = 0.5, color = "red") +
  labs(x = "Travel Time (hours)",
       y = "Probability") +
  theme_minimal()

p3 <- ggplot(time_df, aes(x = 10.67-t, y = likelihood)) +
  geom_area(fill = "salmon", alpha = 0.5, color = "red") +
  labs(x = "Travel Time (hours)",
       y = "Probability") +
  theme_minimal()

p4 <- ggplot(time_df, aes(x = t, y = triangle)) +
  geom_area(fill = "salmon", alpha = 0.5, color = "red") +
  labs(x = "Travel Time (hours)",
       y = "Probability") +
  theme_minimal()
p1
```



The function shows that your net speed is most likely to be 2 mph, but due to fluctuating currents, it could be as low as 1 mph or as high as 3 mph.

1. What's the longest possible duration for the 8-mile trip?

```{mcq}
#| label: fhm-1-udw2
#| inline: true
1. 4 hours
2. 5 hours
3. 6.67 hours
4. 8 hours [correct]
```


2. What's the shortest possible duration for the 8-mile trip?

```{mcq}
#| label: fhm-2-34df
#| inline: true
1. 2 hours
2. 2.67 hours [correct]
3. 3 hours
4. 4 hours
```

3. What's the most likely duration?

```{mcq}
#| label: fhm-3-9t3
#| inline: true
1. 3 hours
2. 4 hours [correct]
3. 5 hours
4. 6.67 hours
```

```{r}
#| label: fig-ftm-3
#| echo: false
#| layout-ncol: 3
#| fig-cap: "Candidate relative probability functions for trip duration."
#| fig-subcap:
#| - " "
#| - " "
#| - " "
p2
p3
p4
```

@fig-ftm-3 shows three different relative probability functions, each purporting to show the distribution of travel times. Only one is correct. Which one?

```{mcq}
#| label: fhm-4-rdf
#| inline: true
1. a [correct]
2. b 
3. c
```





:::
 <!-- end of exr-fly-hide-mattress -->
