---
number-offset: 1
crossref:
  exr-title: "Drill 2."
filters:
  - live
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(drill)
```

# Drill Questions: Systems, dimensions, and data {-}

```{r include=FALSE}
source("_start-up.R")
```

{{< include _devoirs.qmd >}}

`r devoirs::push_answer_style("block")`

:::: {.callout-note collapse=true}
## Systems and dimensions

`r cc <- drill::new_drill_counter()`
`r id <- "d01"`

::: {.panel-tabset}

## `r cc()`

A system is a collection of elements that interact with one another.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

In this book, what technical word is used for the elements in a system? that interact with one another.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`




## `r cc()`

In this book, the word **dimension** refers only to a spatial measurement such as length, width, or height.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="Here, dimension means an aspect, characteristic, or property of a system. It is not restricted to spatial dimensions.")`

## `r cc()`

Which of the following is a dimension of a system in the sense used in this chapter?

```{mcq}
#| label: d01-dimension
#| inline: true
1. A property or aspect of the system [correct]
2. Only a numerical measurement
3. Only a spatial direction
4. Only a column in a data frame
```

## `r cc()`

Which statement best distinguishes a dimension from a quantity?

```{mcq}
#| label: d01-dimension-quantity
#| inline: true
1. A dimension is an aspect of a system; a quantity is a measurable way of representing that aspect. [correct]
2. A quantity is an aspect of a system; a dimension is always a number.
3. Dimension and quantity mean exactly the same thing.
4. A dimension must be measured in meters.
```

## `r cc()`

Which example illustrates **operationalization**?

```{mcq}
#| label: d01-operationalization
#| inline: true
1. Calling anxiety a dimension of personality.
2. Using a blood-pressure measurement as one way to quantify health. [correct]
3. Saying that health is important.
4. Listing several people in a data frame.
```

## `r cc()`

Operationalization is the process of converting a dimension into a quantity that can be measured.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

A good operationalization must be a perfect and complete representation of the dimension it measures.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="The chapter emphasizes that operationalizations can be imperfect. The relevant question is whether an operationalization is good enough for the purpose at hand.")`

## `r cc()`

Why is the heat index used in the chapter as an example of operationalization?

```{mcq}
#| label: d01-heat-index
#| inline: true
1. It provides a quantitative indicator related to comfort in hot, humid weather. [correct]
2. It measures every aspect of human comfort.
3. It is simply another word for temperature.
4. It eliminates the need for a model.
```

:::
::::

:::: {.callout-note collapse=true}
## Data frames and data

`r cc <- drill::new_drill_counter(style = "ALPHA")`

::: {.panel-tabset}

## `r cc()`

In this chapter, a **data frame** is a rectangular array of cells organized into rows and columns.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

In a data frame, each row represents one specimen. Every row should represent a specimens of the same object type.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Which statement about a variable in a data frame is consistent with the terminology used in this book?

```{mcq}
#| label: d01-variable
#| inline: true
1. A variable is a column of a data frame. [correct]
2. A variable is always an unknown number.
3. A variable is always a continuous numerical measurement.
4. A variable is the entire data frame.
```

## "between"

A good test for whether a variable in a data frame should be represented *quantitatively* is that the idea of "between" makes natural sense for values. Keeping this in mind, consider these following situations and decide whether "between" is meaningful.

## `r cc()`

Danish is meaningfully between Chinese and English.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="Although 'D' is alphabetically between 'C' and 'E', this has nothing to do with the nature of the three languages.")`


## `r cc()`

A column containing student ID numbers is necessarily a quantitative variable because its values are written as digits.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="An ID can be a categorical variable even when its values consist of digits. The important feature of 'betweenness' doesn't have meaning for ID numbers.")`

## `r cc()`

Which of the following could be the **object type** of a data frame?

```{mcq}
#| label: d01-object-type
#| inline: true
1. Individual houses for sale [correct]
2. The column named price
3. The number 250
4. The spreadsheet software used to display the data
```

## `r cc()`

In the terminology of this book, what is a **value** in a data frame?

```{mcq}
#| label: d01-value
#| inline: true
1. The entry at the intersection of a variable and a row. [correct]
2. The name of the data frame.
3. The object type of the data frame.
4. The entire column.
```

## `r cc()`

Which statement best describes a **sample**?

```{mcq}
#| label: d01-sample
#| inline: true
1. A smaller collection selected from a larger population. [correct]
2. Any single example of an object.
3. A synonym for a variable.
4. A data frame containing exactly one specimen.
```

## `r cc()`

A population, in the statistical terminology used in this chapter, must consist of people.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="Here, population can refer to any source of specimens, such as people, stars, or events.")`

## `r cc()`

Suppose a data frame contains measurements of 898 adult children, with variables for mother's height, father's height, sex, and the child's height. A smaller data frame containing some of those children can be described as:

```{mcq}
#| label: d01-data-frame-sample
#| inline: true
1. A sample from the larger data frame. [correct]
2. A new object type.
3. A dimension of the original data frame.
4. A conversion factor.
```

:::
::::
