---
number-offset: 1
crossref:
  exr-title: "Drill 1."
filters:
  - live
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(drill)
```

# Drill Questions: Argument and reasoning {-}

```{r include=FALSE}
source("_start-up.R")
```

{{< include _devoirs.qmd >}}

`r devoirs::push_answer_style("block")`

:::: {.callout-note collapse=true}
## Argument and reasoning

`r cc <- drill::new_drill_counter()`
`r id <- "d01"`

::: {.panel-tabset}

## `r cc()`

For purposes of this book, **reasoning** is the transformation of givens into new beliefs and understandings.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Which of the following can be a **given** at the start of a reasoning process?

```{mcq}
#| label: d01-givens
#| inline: true
1. An observation [correct]
2. Only a mathematical equation
3. Only a fact known with certainty
4. Only a numerical measurement
```

## `r cc()`

Hearing a rooster crow is a direct observation that it is dawn.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="The rooster sound is a given. 'It is dawn' is a belief reached by reasoning from the given.")`

## `r cc()`

What links the observation of a rooster sound to the reasoned belief that it is dawn?

```{mcq}
#| label: d01-rooster-link
#| inline: true
1. The heuristic rule that roosters crow at dawn. [correct]
2. The fact that dawn always occurs when a rooster makes a sound.
3. A formal proof that dawn has occurred.
4. A measurement of the time of day.
```

## `r cc()`

A **heuristic** is a reasoning link that is useful but is not guaranteed to be correct.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Which statement best describes **formal reasoning** as used in this chapter?

```{mcq}
#| label: d01-formal
#| inline: true
1. It uses links governed by rules that are accepted as certain. [correct]
2. It is reasoning based only on common sense.
3. It is always about real-world situations.
4. It cannot contain human errors.
```

## `r cc()`

A premise is a statement assumed to be true for the purpose of formal reasoning.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Why can formal reasoning give misleading results when it is applied to a real-world situation?

```{mcq}
#| label: d01-premises
#| inline: true
1. The premises used in the formal reasoning may not accurately represent the real-world situation. [correct]
2. Formal reasoning is never reliable.
3. Formal reasoning always uses numerical data.
4. Formal reasoning cannot be carried out by a computer.
```

:::
::::

:::: {.callout-note collapse=true}
## Models as bridges

`r cc <- drill::new_drill_counter(style = "ALPHA")`

::: {.panel-tabset}

## `r cc()`

A model is a **representation** of an item, process, or setting of interest.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Which statement best expresses the role of a quantitative model in this book?

```{mcq}
#| label: d01-model-bridge
#| inline: true
1. It bridges real-world givens and the power of formal reasoning. [correct]
2. It replaces the real world with a mathematically exact copy.
3. It guarantees that conclusions about the real world are true.
4. It eliminates the need for heuristic reasoning.
```

## `r cc()`

Building a quantitative model is itself an exercise in formal reasoning.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="Model construction uses knowledge, understanding, and belief about real-world matters together with heuristic reasoning. The completed model can then serve as a premise for formal reasoning.")`

## `r cc()`

Suppose a model is used to predict the effect of an intervention in a real-world situation, for instance, a medical treatment. The model produces a logically correct result, but the model itself leaves out an important feature of the real situation. Which conclusion is most consistent with the chapter?

```{mcq}
#| label: d01-model-fault
#| inline: true
1. The formal calculation can be correct even though the conclusion is not useful for the real-world situation. [correct]
2. The formal calculation must contain an arithmetic error.
3. The model must therefore be a fact rather than a representation.
4. Formal reasoning cannot be used with models.
```

## `r cc()`

The usefulness of a model depends in part on the purpose for which the model is built.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

Which statement best describes the iterative nature of modeling?

```{mcq}
#| label: d01-iterative
#| inline: true
1. Build a model, examine how well it captures important aspects of reality, and improve it when needed. [correct]
2. Build a model once and then treat it as permanently correct.
3. Replace heuristic reasoning with formal reasoning before building the model.
4. Avoid checking a model against reality because models are representations.
```

## `r cc()`

In the epidemic example, the SIR model is used to represent the connections among **susceptible**, **infective**, and **recovered** people.

`r devoirs::devoirsTF(correct = TRUE, id = cc(id))`

## `r cc()`

In the epidemic example, the computer simulation is itself the model; the SIR model is merely the software used to run it.

`r devoirs::devoirsTF(correct = FALSE, id = cc(id), wrong.hint="The SIR description is the model. The computer program implements the model's bookkeeping and formal calculations.")`

:::
::::
