<!--
id: "deer-spend-roof"
created: "Thu Aug 28 16:07:34 2025"
attribution: TBA
status: "draft"
chapter: 14
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}

::: {#exr-deer-spend-roof}
Writing all numbers in scientific notation, as $1 \times 10^{x}$.

Consider the number $3.814697 \time 10^{-5}$. This is between 1e-4 and 1e-5, as the following calculation demonstrates.

```{r}
#| label: C14-fueed
mynum <- 3.814697e-5
mynum < 1e-4
mynum > 1e-5
```

I want to be able to write `mynum` in a format where the mantissa is *always* 1, but where the exponent might not be an integer. For instance, in the example, `mynum` is equal to $1 \times 10^{-4.41854}$. I'll call the exponent when the mantissa is 1 the **magnitude** of the number.

The magnitude of 10 is 1, the magnitude of 1 is 0, the magnitude of 0.1 is -1, and so on.

Given two numbers, the magnitude of their product is the *sum* of the magnitudes of the two numbers. Example: 10 has magnitude 1 and 0.01 has magnitude -2, so $10 \times 0.01 = 0.1$ has magnitude $1 - 2 = -1$.



```{r}
#| label: C14-qkscl
Magnitudes <- tibble::tibble(
  mag = 1 / (2^(0:10)) 
)

# initialize using square roots
num <- rep(10, nrow(Magnitudes))
for (k in 2:length(num)) num[k] <- sqrt(num[k-1])
Magnitudes$num <- num

augment <- function(mag_table) {
  Another <- tibble::tibble(
    mag = cumsum(mag_table[-1, ]$mag),
    num = cumprod(mag_table[-1, ]$num)
  )
  mag_table <- dplyr::bind_rows(mag_table, Another)
  And_another <- mag_table |>
    dplyr::mutate(num = sqrt(num), mag = mag/2)
  
  dplyr::bind_rows(mag_table, And_another) |>
    clean_tab()
  
}

pick_some <- function(mag_table, n = nrow(mag_table), pickn = 2L) {
  result <- tibble::tibble(mag = numeric(n), num = numeric(n))
  for (k in 1:n) {
    rows <- base::sample.int(nrow(mag_table), size = pickn)
    tmp <- mag_table[rows, ]
    result[k, "num"] <- prod(tmp$num)
    result[k, "mag"] <- sum(tmp$mag)
  }
  
  dplyr::bind_rows(mag_table, result) |> clean_tab() |> unique()
}

promote <- function(mag_table) {
  newtable <- mag_table |> 
    dplyr::mutate(
      mag = mag + 1/2,
      num = num * sqrt(10)
    ) 
  
  dplyr::bind_rows(mag_table, newtable) |> clean_tab()
}

# get rid of anything bigger than 10
clean_tab <- function(mag_table) {
  tmp <- 
    mag_table |>
    dplyr::filter(num < 100) |>
    dplyr::mutate(mag = ifelse(num > 10, mag-1, mag),
                  num = ifelse(num > 10, num/10, num)) |>
    dplyr::arrange(desc(num))
  kill <- which(abs(diff(tmp$num)) < 1e-12)
  tmp[-kill,]
}

mtable <- Magnitudes
for (k in 1:3) {
  mtable <- augment(promote(mtable))
  mtable <- pick_some(promote(mtable), pickn = 2)
  mtable <- promote(mtable)
}
```

Now we have a large number of points for log10(x) in the domain 1 <= x <= 10

We can construct extensions by multiplying or dividing the number by 10, adding or subtracting 1 respectively from the mag.
:::
 <!-- end of exr-deer-spend-roof -->
