<!--
id: "maple-shut-pantry"
created: "Mon Dec 22 13:15:23 2025"
attribution: TBA
status: "ready-to-use"
chapter: 4
mode: "exercise"
-->

{{< include ../exercise-header.qmd >}}



::: {#exr-maple-shut-pantry}
Consider the data on the body mass (kg) and brain mass (gm) for 62 species of mammals shown in @fig-mammal-mass. (You can hover the cursor over a point to see the species name.)

```{r}
#| echo: false
#| label: fig-mammal-mass
#| warning: false
library(ggformula)
library(ggiraph)
mammals <- read.csv("../www/mammals.csv")
P <- mammals |> 
  ggformula:::gf_point_interactive(brain_weight ~ body_weight,
                       tooltip = paste("Species: ", mammals$species)) |>
  gf_refine(
    scale_y_log10(),
    scale_x_log10(),
    coord_fixed()
  ) |>
  gf_lm(interval = "prediction", level = 0.67) |>
  gf_labs(y = "Brain mass (gm)", x = "Body mass (kg)") |>
  gf_theme(theme_minimal())
P |> gf_girafe()
```

1. The lesser short-tailed shrew has both the smallest body mass and brain mass. What is the brain mass of the short-tailed shrew? (Pick the answer that's closest.)

```{mcq}
#| label: msp-1-4lse
#| inline: true
1. 0.03 gm
2. 0.14 gm [correct]
3. 0.52 gm
```

2. The water opposum, arctic fox, nine-banded armadillo, and rock hyrax (in the middle of the cloud of points) all have about the same body mass. What is that body mass? (Pick the answer that's closest.) 

```{mcq}
#| label: msp-2-sddo
#| inline: true
1. 1-2 kg
2. 3-4 kg [correct]
3. 6-7 kg
4. 8-9 kg
```

3. Which of these animals has a body mass closest to 10 kg?

```{mcq}
#| label: msp-3-fhsl
#| inline: true
1. raccoon
1. baboon [correct]
2. roe deer
4. vervet
```

4. Of the mammals in the range of body masses from 10 to 100 kg, which has the smallest brain size?

```{mcq}
#| label: msp-4-le4
#| inline: true
1. water opposum [hint: It's smaller than 10 kg, so not the answer to the question.]
2. giant armadillo
3. goat
4. kangaroo [correct]
```

5. The gray bar surrounding the blue line in @fig-mammal-mass contains about 2/3 of the 62 mammals plotted. Which animal has the smallest ratio of brain mass compared its range of body masses?

```{mcq}
#| label: msp-5-odr
#| inline: true
1. water opposum [correct]
2. tenrec
3. musk shrew
4. desert hedgehog
```

@fig-mammal-sleep shows the daily sleeping time versus lifetime for the 62 mammals. The level of danger from predation that the animal is exposed to is shown by color.

```{r}
#| echo: false
#| label: fig-mammal-sleep
#| warning: false
library(ggformula)
library(ggiraph)
mammals <- read.csv("../www/mammals.csv")
P <- mammals |> 
  ggformula:::gf_point_interactive(total_sleep ~ lifespan, color = ~ danger,
                       tooltip = paste("Species: ", mammals$species, "<br>Danger", mammals$danger)) |>
  gf_refine(
    scale_x_log10(),
    scale_color_viridis_c()
  ) |>
  gf_labs(y = "Sleep (hrs per day)", x = "Lifespan (yr)") |>
  gf_theme(theme_minimal())
P |> gf_girafe()
```

6. The mammal who sleeps the least sleeps how many hours per day?

```{mcq}
#| label: msp-6-kew
1. 2.5 [correct]
2. 2.9
3. 3.5
4. Can't tell the difference on this scale.
```

7. A chinchilla lives about 7 years. What level of danger does the chinchilla face?

```{mcq}
#| label: msp-7-ese
#| inline: true
1. 1
2. 2
3. 3
4. 4 [correct]
5. 5
```


8. The Sleep axis is linear, while the Lifespan axis is magnitude. Explain why this choice makes sense (or if it doesn't).

`r devoirs::devoirs_text("msp-8-klj")`

:::
 <!-- end of exr-maple-shut-pantry -->
